---
title: "new_project_visa"
output: html_document
---

#libraries
```{r}
library(randomForest)
library(party)
library(dplyr)
library(tidyr)
library(sqldf)
library(caret)
library(glmnet)
library(car)
library(ROCR)
library(e1071)
library(gbm)
library(class)
library(MASS)
library(readr)
library(tree)
library(readr)

```

#Load Data set
```{r}
h1b_kaggle <- read.csv("h1b_kaggle.csv", header=TRUE, stringsAsFactors = FALSE)
```


#Preprocessing data
```{r}
#find rows where conditions are true where CASE_STATUS is CERTIFIED, CERTIFIED-WITHDRAWN that is treated as CERTIFIED or DENIED in the year 2014, 2015 and 2016

myData = filter(h1b_kaggle, h1b_kaggle$CASE_STATUS %in% c('CERTIFIED','DENIED', 'CERTIFIED-WITHDRAWN')  & (h1b_kaggle$YEAR == 2016 | h1b_kaggle$YEAR == 2015 | h1b_kaggle$YEAR == 2014))


#Keep only complete cases
myData = myData[complete.cases(myData),]

h1bData = myData

#Eliminate columns case#, employer, job title, long, lat
h1bData[,c(1,3,5,10,11)]=NULL

#Create a new column called worksite to keep only state
h1bData=separate(data = h1bData, col = WORKSITE, into = c("CITY", "STATE"), sep = ",")

#Create a new column to save occupations
h1bData$occ=NA

#Keep occupations containing the keyword and set the new occupation
h1bData$occ[grep("engineer",h1bData$SOC_NAME, ignore.case = T)]="ENGINEER"
h1bData$occ[grep("manager",h1bData$SOC_NAME, ignore.case = T)]="MANAGER"
h1bData$occ[grep("technician",h1bData$SOC_NAME, ignore.case = T)]="TECHNICIAN"
h1bData$occ[grep("teacher",h1bData$SOC_NAME, ignore.case = T)]="TEACHER"
h1bData$occ[grep("executive",h1bData$SOC_NAME, ignore.case = T)]="EXECUTIVE"
h1bData$occ[grep("accountant",h1bData$SOC_NAME, ignore.case = T)]="ACCOUNTANT"
h1bData$occ[grep("actor",h1bData$SOC_NAME, ignore.case = T)]="ACTOR"
h1bData$occ[grep("advertising",h1bData$SOC_NAME, ignore.case = T)]="ADVERTISING"
h1bData$occ[grep("lawyer",h1bData$SOC_NAME, ignore.case = T)]="LAWYER"
h1bData$occ[grep("financial",h1bData$SOC_NAME, ignore.case = T)]="FINANCIAL"
h1bData$occ[grep("arquitect",h1bData$SOC_NAME, ignore.case = T)]="ARQUITECT"
h1bData$occ[grep("programmer",h1bData$SOC_NAME, ignore.case = T)]="SOFTWARE"
h1bData$occ[grep("software",h1bData$SOC_NAME, ignore.case = T)]="SOFTWARE"
h1bData$occ[grep("computer",h1bData$SOC_NAME, ignore.case = T)]="SOFTWARE"
h1bData$occ[grep("developer",h1bData$SOC_NAME, ignore.case = T)]="SOFTWARE"
h1bData$occ[grep("analyst",h1bData$SOC_NAME, ignore.case = T)]="ANALYST"
h1bData$occ[grep("scien",h1bData$SOC_NAME, ignore.case = T)]="SCIENTIST"
h1bData$occ[grep("specialist",h1bData$SOC_NAME, ignore.case = T)]="SPECIALIST"
h1bData$occ[grep("animal",h1bData$SOC_NAME, ignore.case = T)]="ANIMAL RELATED"
h1bData$occ[grep("athlet",h1bData$SOC_NAME, ignore.case = T)]="ATHLETE"
h1bData$occ[grep("cook",h1bData$SOC_NAME, ignore.case = T)]="COOK"
h1bData$occ[grep("chef",h1bData$SOC_NAME, ignore.case = T)]="COOK"
h1bData$occ[grep("admin",h1bData$SOC_NAME, ignore.case = T)]="ADMINISTRATIVE"

#Eliminate columns SOC_NAME and CITY
h1bData$SOC_NAME=NULL
h1bData$CITY= NULL

#Removing states with low count
a=sqldf("select count(*) cc, STATE from 'h1bData' group by STATE")      
b=sqldf("select * from a where cc>2000 AND STATE <> ' NA'")
h1bData$STATE=ifelse(h1bData$STATE %in% b$STATE,h1bData$STATE,NA)         

#Convert the dependent variable to binary
h1bData$CASE_STATUS=ifelse(h1bData$CASE_STATUS %in% c("CERTIFIED-WITHDRAWN", "CERTIFIED"),"1","0")

#Converting categorical variables into factors
h1bData[,c(-3)]= lapply(h1bData[,c(-3)], as.factor)
h1bData = h1bData[complete.cases(h1bData),]

#Using years 2014 and 2015 as training and 2016 as test
data.test = h1bData[0:557090, ]
data.train = h1bData[557091:1522982,]
data.test = data.test[, -4]
data.train = data.train[, -4]
```


#LOGISTIC REGRESSION USING ALL PREDICTORS
```{r}
#Fitting the model on the training dataset
h1bglm.train.fit =glm(CASE_STATUS~., family=binomial(link = logit), data = data.train)

#Finding Prdicitons on Testing set
prediction=predict(h1bglm.train.fit, newdata=data.test, type="response")
View(prediction)

#Threshold
pred_class = vector()
pred_class[prediction<0.5]=0
pred_class[prediction>=0.5]=1

#confusion matrix
confusionMatrix(pred_class, data.test$CASE_STATUS,positive="1")
#levels(data.test$CASE_STATUS)

#ROC Curve
pred = prediction(pred_class, data.test$CASE_STATUS)
perf = performance(pred,"tpr","fpr")
plot(perf)

#Area Under the Curve
auc.tmp = performance(pred,"auc");
auc = as.numeric(auc.tmp@y.values)
auc
```

#LOGISTIC REGRESSION USING ONLY FULL_TIME_POSITION AND PREVAILING_WAGE
```{r}
#Fitting the model on the training dataset
h1bglm.train.fit =glm(CASE_STATUS~.-STATE-occ, family=binomial(link = logit), data = data.train)

#Finding Prdicitons on Testing set
prediction=predict(h1bglm.train.fit, newdata=data.test, type="response")

#Threshold
pred_class = vector()
pred_class[prediction<0.5]=0
pred_class[prediction>=0.5]=1

#confusion matrix
confusionMatrix(pred_class, data.test$CASE_STATUS,positive="1")

#ROC Curve
pred = prediction(pred_class, data.test$CASE_STATUS)
perf = performance(pred,"tpr","fpr")
plot(perf)

#Area Under the Curve
auc.tmp = performance(pred,"auc");
auc = as.numeric(auc.tmp@y.values)
auc
```

#LOGISTIC REGRESSION USING ONLY FULL_TIME_POSITION, PREVAILING_WAGE AND STATE
```{r}
#Fitting the model on the training dataset
h1bglm.train.fit =glm(CASE_STATUS~.-occ, family=binomial(link = logit), data = data.train)

#Finding Prdicitons on Testing set
prediction=predict(h1bglm.train.fit, newdata=data.test, type="response")

#Threshold
pred_class = vector()
pred_class[prediction<0.5]=0
pred_class[prediction>=0.5]=1

#confusion matrix
confusionMatrix(pred_class, data.test$CASE_STATUS,positive="1")

#ROC Curve
pred = prediction(pred_class, data.test$CASE_STATUS)
perf = performance(pred,"tpr","fpr")
plot(perf)

#Area Under the Curve
auc.tmp = performance(pred,"auc");
auc = as.numeric(auc.tmp@y.values)
auc
```

#LOGISTIC REGRESSION USING ONLY FULL_TIME_POSITION, PREVAILING_WAGE AND OCCUPATION
```{r}
#Fitting the model on the training dataset
h1bglm.train.fit =glm(CASE_STATUS~.-STATE, family=binomial(link = logit), data = data.train)

#Finding Prdicitons on Testing set
prediction=predict(h1bglm.train.fit, newdata=data.test, type="response")

#Threshold
pred_class = vector()
pred_class[prediction<0.5]=0
pred_class[prediction>=0.5]=1

#confusion matrix
confusionMatrix(pred_class, data.test$CASE_STATUS,positive="1")

#ROC Curve
pred = prediction(pred_class, data.test$CASE_STATUS)
perf = performance(pred,"tpr","fpr")
plot(perf)

#Area Under the Curve
auc.tmp = performance(pred,"auc");
auc = as.numeric(auc.tmp@y.values)
auc
```

#LDA USING USING ALL PREDICTORS
```{r}
lda.fit=lda(CASE_STATUS~., data=data.train)
lda.predict = predict(lda.fit, data.test, type = "prob")
confusionMatrix(lda.predict$class, data.test$CASE_STATUS,positive="1")
```

#LDA USING ONLY FULL_TIME_POSITION AND PREVAILING_WAGE
```{r}
lda.fit=lda(CASE_STATUS~.-STATE-occ, data=data.train)
lda.predict = predict(lda.fit, data.test, type = "prob")
confusionMatrix(lda.predict$class, data.test$CASE_STATUS,positive="1")
```

#LDA USING ONLY FULL_TIME_POSITION, PREVAILING_WAGE AND STATE
```{r}
lda.fit=lda(CASE_STATUS~.-STATE, data=data.train)
lda.predict = predict(lda.fit, data.test, type = "prob")
confusionMatrix(lda.predict$class, data.test$CASE_STATUS,positive="1")
```

#LDA USING ONLY FULL_TIME_POSITION, PREVAILING_WAGE AND OCCUPATION
```{r}
lda.fit=lda(CASE_STATUS~.-occ, data=data.train)
lda.predict = predict(lda.fit, data.test, type = "prob")
confusionMatrix(lda.predict$class, data.test$CASE_STATUS,positive="1")
```


#QDA USING USING ALL PREDICTORS
```{r}
qda.fit=qda(CASE_STATUS~., data=data.train)
qda.predict = predict(qda.fit, data.test)
View(qda.predict)
confusionMatrix(qda.predict$class, data.test$CASE_STATUS,positive="1")
```

#QDA USING ONLY FULL_TIME_POSITION AND PREVAILING_WAGE
```{r}
lda.fit=lda(CASE_STATUS~.-STATE-occ, data=data.train)
lda.predict = predict(lda.fit, data.test, type = "prob")
confusionMatrix(lda.predict$class, data.test$CASE_STATUS,positive="1")
```

#QDA USING ONLY FULL_TIME_POSITION, PREVAILING_WAGE AND STATE
```{r}
lda.fit=lda(CASE_STATUS~.-STATE, data=data.train)
lda.predict = predict(lda.fit, data.test, type = "prob")
confusionMatrix(lda.predict$class, data.test$CASE_STATUS,positive="1")
```

#QDA USING ONLY FULL_TIME_POSITION, PREVAILING_WAGE AND OCCUPATION
```{r}
lda.fit=lda(CASE_STATUS~.-occ, data=data.train)
lda.predict = predict(lda.fit, data.test, type = "prob")
confusionMatrix(lda.predict$class, data.test$CASE_STATUS,positive="1")
```

#LOGISTIC REGRESSION USING ALL PREDICTORS and threshold 0.97
```{r}
#Fitting the model on the training dataset
h1bglm.train.fit =glm(CASE_STATUS~., family=binomial(link = logit), data = data.train)

#Finding Predicitons on Testing set
prediction=predict(h1bglm.train.fit, newdata=data.test, type="response")

#Threshold
pred_class = vector()
pred_class[prediction<0.97]=0
pred_class[prediction>=0.97]=1

##confusion matrix
confusionMatrix(pred_class, data.test$CASE_STATUS,positive="1")

#ROC Curve
pred = prediction(pred_class, data.test$CASE_STATUS)
perf = performance(pred,"tpr","fpr")
plot(perf)

#Area Under the Curve
auc.tmp = performance(pred,"auc");
auc = as.numeric(auc.tmp@y.values)
auc
```

#LOGISTIC REGRESSION USING ONLY FULL_TIME_POSITION AND PREVAILING_WAGE
```{r}
#Fitting the model on the training dataset
h1bglm.train.fit =glm(CASE_STATUS~.-STATE-occ, family=binomial(link = logit), data = data.train)

#Finding Prdicitons on Testing set
prediction=predict(h1bglm.train.fit, newdata=data.test, type="response")

#Threshold
pred_class = vector()
pred_class[prediction<0.97]=0
pred_class[prediction>=0.97]=1

##confusion matrix
confusionMatrix(pred_class, data.test$CASE_STATUS,positive="1")

#ROC Curve
pred = prediction(pred_class, data.test$CASE_STATUS)
perf = performance(pred,"tpr","fpr")
plot(perf)

#Area Under the Curve
auc.tmp = performance(pred,"auc");
auc = as.numeric(auc.tmp@y.values)
auc
```

#LOGISTIC REGRESSION USING ONLY FULL_TIME_POSITION, PREVAILING_WAGE AND STATE
```{r}
#Fitting the model on the training dataset
h1bglm.train.fit =glm(CASE_STATUS~.-occ, family=binomial(link = logit), data = data.train)

#Finding Prdicitons on Testing set
prediction=predict(h1bglm.train.fit, newdata=data.test, type="response")

#Threshold
pred_class = vector()
pred_class[prediction<0.97]=0
pred_class[prediction>=0.97]=1

##confusion matrix
confusionMatrix(pred_class, data.test$CASE_STATUS,positive="1")

#ROC Curve
pred = prediction(pred_class, data.test$CASE_STATUS)
perf = performance(pred,"tpr","fpr")
plot(perf)

#Area Under the Curve
auc.tmp = performance(pred,"auc");
auc = as.numeric(auc.tmp@y.values)
auc
```

#LOGISTIC REGRESSION USING ONLY FULL_TIME_POSITION, PREVAILING_WAGE AND OCCUPATION
```{r}
#Fitting the model on the training dataset
h1bglm.train.fit =glm(CASE_STATUS~.-STATE, family=binomial(link = logit), data = data.train)

#Finding Prdicitons on Testing set
prediction=predict(h1bglm.train.fit, newdata=data.test, type="response")

#Threshold
pred_class = vector()
pred_class[prediction<0.97]=0
pred_class[prediction>=0.97]=1

#confusion matrix
confusionMatrix(pred_class, data.test$CASE_STATUS,positive="1")

#ROC Curve
pred = prediction(pred_class, data.test$CASE_STATUS)
perf = performance(pred,"tpr","fpr")
plot(perf)

#Area Under the Curve
auc.tmp = performance(pred,"auc");
auc = as.numeric(auc.tmp@y.values)
auc
```

#DECISION TREE USING ALL PREDICTORS AND THRESHOLD 0.97
```{r}
ctree1=ctree(CASE_STATUS~., data=data.train)
probabilities = 1-unlist(treeresponse(ctree1,newdata=data.test), use.names=F)[seq(1,nrow(data.test)*2,2)]

probabilities[probabilities<0.97]=0
probabilities[probabilities>=0.97]=1

confusionMatrix(probabilities, data.test$CASE_STATUS,positive="1")

#ROC CURVE
pred1 = prediction( probabilities, data.test$CASE_STATUS)
perf1 = performance(pred1,"tpr","fpr")
plot(perf1)

#AREA Under the Curve
auc.tmp1 = performance(pred1,"auc");
auc1 = as.numeric(auc.tmp1@y.values)
auc1
```
#DECISION TREE USING ALL PREDICTORS AND THRESHOLD 0.5
```{r}
ctree1=ctree(CASE_STATUS~., data=data.train)
probabilities = 1-unlist(treeresponse(ctree1,newdata=data.test), use.names=F)[seq(1,nrow(data.test)*2,2)]

probabilities[probabilities<0.5]=0
probabilities[probabilities>=0.5]=1

confusionMatrix(probabilities, data.test$CASE_STATUS,positive="1")

#ROC CURVE
pred1 = prediction( probabilities, data.test$CASE_STATUS)
perf1 = performance(pred1,"tpr","fpr")
plot(perf1)

#AREA Under the Curve
auc.tmp1 = performance(pred1,"auc");
auc1 = as.numeric(auc.tmp1@y.values)
auc1
```


#DECISION TREE USING ONLY FULL_TIME_POSITION AND PREVAILING_WAGE
```{r}
ctree1=ctree(CASE_STATUS~.-STATE-occ, data=data.train)
probabilities = 1-unlist(treeresponse(ctree1,newdata=data.test), use.names=F)[seq(1,nrow(data.test)*2,2)]

probabilities[probabilities<0.97]=0
probabilities[probabilities>=0.97]=1

confusionMatrix(probabilities, data.test$CASE_STATUS,positive="1")

#ROC CURVE
pred1 = prediction( probabilities, data.test$CASE_STATUS)
perf1 = performance(pred1,"tpr","fpr")
plot(perf1)

#AREA Under the Curve
auc.tmp1 = performance(pred1,"auc");
auc1 = as.numeric(auc.tmp1@y.values)
auc1
```

#DECISION TREE USING ONLY FULL_TIME_POSITION, PREVAILING_WAGE AND STATE
```{r}
ctree1=ctree(CASE_STATUS~.-occ, data=data.train)
probabilities = 1-unlist(treeresponse(ctree1,newdata=data.test), use.names=F)[seq(1,nrow(data.test)*2,2)]

probabilities[probabilities<0.97]=0
probabilities[probabilities>=0.97]=1

confusionMatrix(probabilities, data.test$CASE_STATUS,positive="1")

#ROC CURVE
pred1 = prediction( probabilities, data.test$CASE_STATUS)
perf1 = performance(pred1,"tpr","fpr")
plot(perf1)

#AREA Under the Curve
auc.tmp1 = performance(pred1,"auc");
auc1 = as.numeric(auc.tmp1@y.values)
auc1
```

#DECISION TREE USING ONLY FULL_TIME_POSITION, PREVAILING_WAGE AND OCCUPATION
```{r}
ctree1=ctree(CASE_STATUS~.-STATE, data=data.train)
probabilities = 1-unlist(treeresponse(ctree1,newdata=data.test), use.names=F)[seq(1,nrow(data.test)*2,2)]

probabilities[probabilities<0.97]=0
probabilities[probabilities>=0.97]=1

confusionMatrix(probabilities, data.test$CASE_STATUS,positive="1")

#ROC CURVE
pred1 = prediction( probabilities, data.test$CASE_STATUS)
perf1 = performance(pred1,"tpr","fpr")
plot(perf1)

#AREA Under the Curve
auc.tmp1 = performance(pred1,"auc");
auc1 = as.numeric(auc.tmp1@y.values)
auc1
```

